Intrinsic Verification of Parsers and Formal Grammar Theory in Dependent Lambek Calculus

### 2024-11-13 paper/paper.tex
start editing section 4. Done for the night

The syntax of
\theoryabbv is based on the dependent linear-non-linear (\lnld)
calculus from \cite{krishnaswami_integrating_2015}. As in \lnld, \theoryabbv includes both
non-linear dependent types, as well as linear types, which are allowed
to depend on the non-linear types, but not on other linear types.
%
The main point of departure from \lnld's design in that the linear
typing is \emph{non-commutative}, i.e., that exchange is not an
admissible structural rule. Furthermore, we add a general-purpose
indexed inductive linear type connective, as well as an
\emph{equalizer} type, which we will show allows us to perform
inductive proofs of equalities between linear terms.
%
Finally, we add base types and axioms to the system specifically to
model formal grammars and parsing.

@@

The non-linear types of \theoryabbv include the standard dependent
types for $\Pi,\Sigma$, extensional equality, natural numbers,
booleans, unit and empty types. We include these rules in the
appendix. We present the other non-linear type constructions in
\Cref{fig:non-linear-stuff}. First, we include universe types $U$ of
small non-linear types and $L$ of linear types. These are defined as
``universes ala Coquand'' in that we define a judgment saying when a
non-linear type is small and define the universe to internalize
precisely this judgment. These universe types are needed so that we
can define types by recursion on natural numbers. Next, we include a
non-linear type $\ltonl A$ where $A$ is a linear type. The intuition
for this type is that its elements are the linear terms that are
``resource free'': its introduction rule says we can construct an
$\ltonl A$ when we have a linear term of type $A$ with no free linear
variables. This type is used extensively in our examples, playing a
similar role to the $!$ modality of ordinary linear logic.
\pedro{We should also reference the persistence modality $\square$
from separation logic}

>>>>

The design of \theoryabbv is based the dependent
linear-non-linear calculus (\lnld) and Lambek calculus, also known as
non-commutative linear logic.
\cite{krishnaswami_integrating_2015,lambekCalculus}. As in \lnld,
\theoryabbv includes both non-linear dependent types, as well as
linear types, which are allowed to depend on the non-linear types, but
not on other linear types.
%
The main point of departure from \lnld's design is that, as in Lambek calculus \cite{lambekCalculus, the linear
typing is \emph{non-commutative}, i.e., that exchange is not an
admissible structural rule. Furthermore, we add a general-purpose
indexed inductive linear type connective, as well as an
\emph{equalizer} type, which we will show allows us to perform
inductive proofs of equalities between linear terms.
%
Finally, while \lnld was enhanced with special connectives
inspired by separation logic to model imperative programming, we
instead add base types and axioms to the system specifically to model
formal grammars and parsing.

@@

The non-linear types of \theoryabbv include the standard dependent
types for $\Pi,\Sigma$, extensional equality, natural numbers,
booleans, unit and empty types. We include these rules in the
appendix. We present the other non-linear type constructions in
\Cref{fig:non-linear-stuff}. First, we include universe types $U$ of
small non-linear types and $L$ of linear types. These are defined as
universes ``ala Coquand'' in that we define a judgment saying when a
non-linear type is small and define the universe to internalize
precisely this judgment \cite{universesAlaCoquand}. These universe
types are needed so that we can define types by recursion on natural
numbers. Next, we include a non-linear type $\ltonl A$ where $A$ is a
linear type. The intuition for this type is that its elements are the
linear terms that are ``resource free'': its introduction rule says we
can construct an $\ltonl A$ when we have a linear term of type $A$
with no free linear variables. This type is used extensively in our
examples, playing a similar role to the $!$ modality of ordinary
linear logic or the persistence modality $\square$ of separation logic
\cite{linearlogic,persistencemodalityseparationlogic}.
